### 2013-12-12 mcmc.tex
functional testing

Just as most (though not all) MCMC methods require a choice of proposal distribution,
  most (though not all) require a choice about initialization:
The walker needs (or walkers need) to be started somewhere in the parameter space.
In many use cases for MCMC,
  the investigator wants to use it to obtain uncertainty information about
  or propagate uncertainty into parameter estimates.
In these cases, it makes sense to initialize the walker or walkers at sensible parameter estimates,
  found by optimizing a likelihood or a posterior PDF in advance of sampling.
In extremely high dimensions (large numbers of parameters), this can be a bad idea,
  since the optimal parameters are not necessarily near typical posterior samples,
  but in low dimensions (few to tens) this is often sensible.

@@

\paragraph{functional testing:}
...Run on standard distributons.

...Set likelihood to 1 and check that you sample your prior pdf.

>>>>

Just as most (though not all) MCMC methods require a choice of proposal distribution,
  most (though not all) require a choice about initialization:
The walker needs (or walkers need) to be started somewhere in the parameter space.
In many use cases for MCMC,
  the investigator wants to use it to obtain uncertainty information about
  or propagate uncertainty into parameter estimates.
In these cases, it makes sense to initialize the walker or walkers at sensible parameter estimates,
  found by optimizing a likelihood or a posterior PDF in advance of sampling.
In extremely high dimensions (large numbers of parameters), this can be a bad idea,
  since the optimal parameters are not necessarily near typical posterior samples\note{Make the counterintuitive point.},
  but in low dimensions (few to tens) this is often sensible.

@@

When issues arise with MCMC sampling,
  they are sometimes difficult to diagnose:
Is it the MCMC code, the priors, the likelihood function,
  the initialization, the tuning, or something else?

\paragraph{functional testing:}
In addition to a full set of unit tests for every part of your code
  (yeah right!\note{If you aren't unit testing, you are probably making some very big mistakes.}),
  we recommend building some end-to-end functional tests that permit you to perform MCMC
  runs with output that must meet certain expectations.
The simplest kind of functional test is to sample a known distribution,
  and check that the sampling has the moments you expect (to within tolerances).
For example, you can sample a $d$-dimensional Gaussian distribution with known non-trivial covariance,
  and check that the empirical covariance comes out as expected.
This is an easy test to fail, so a success builds confidence in any MCMC code.

In the most extreme form of this test,
  for a Gaussian in $d$ dimensions,
  build the covariance matrix input to the Gaussian distribution
  by taking the sum of the outer product of $d+1$ randomly generated vectors.
Then, before sampling, get the eigenvalues and eigenvectors of the matrix.
Set up the MCMC sampler with the Gaussian function with this covariance matrix
  as the probability function (or the log of it as the log-probability function).
Sample for many autocorrelation times,
  or until you are confident you are getting good samples.
Check that the empirical $d$-dimensional covariance of the samples
  has the same (or similar) eigenvectors and eigenvlues to the input matrix.
Also check that the individual one-dimensional samplings
  (along each axis or any projection in the $d$-space)
  look Gaussian and have nearly vanishing skew and kurtosis.
It takes some thinking and some calculating to see ``how well'' these things all ought to match
  expectations, but one thing that can be said with confidence is that agreement should improve
  as the sampling proceeds.
If it doesn't, it is likely that there is something broken about the sampler itself.

...Set likelihood to 1 and check that you sample your prior pdf.
...Checks properness of priors.
...Checks for think-os about parameterization.
...Make triangle plots.
